raptr: Representative and Adequate Prioritization Toolkit in R

### 2016-11-27 code/rmarkdown/text.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across this study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 8a). The second species was simulated to represent a species with an idealized distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8c). The third species represents a species with two distinct populations (Fig. 2c). This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8d). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 8d} \\
%
\end{align*}

We generated four solutions for each species. First to represent solutions generated using conventional conservation planning methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. Second to show how the addition of geographic targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets. Third to represent solutions using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`. Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.

@@

The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig. 3d). This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented. This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig. 3c) contained the same number of planning units as the solution with just the amount targets (Fig. 3a). Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.

@@

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is highly fragmented.

@@

The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig. 2c). The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig. 3i). While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig. 3k). However, the addition of geographic space targets resolved these issues.

@@

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the realized niche for several species. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63]. We selected Queensland, Australia as the study area. This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterize the environmental variation across the study area (Fig. 4).

We selected four bird species that span a range of different ecological requirements, evolutionary histories, distributions: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_). We then mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig. S2). We used the following method to ensure that the demand points reflected the core parts of the species' realized niches. For each species, we generated random points inside the species range and at each point extracted the principal component values at that location. We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231]. We then generated uniformly distributed points inside the species' kernels and estimated the density of the training points at the uniformly generated points. These uniformly distributed points and associated density estimated were used as demand point coordinates and weights (respectively).

We generated two solutions. The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species. The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.

### Results
Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches--with `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'one')` excpetion. Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`. Yet it failed to achieve this for `r parsed.not.representative.space.held.names`. This result demonstrates that though solutions generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritization securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We calculated the average of the values associated with individuals in each grid cell. These values were used to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].

The grid cells were used as planning units. The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit. Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points. To ensure that the solutions did not prioritize particularly costly areas, we included opportunity cost data (Fig. 7b). We obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.

@@

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs. 8a and 8c). This solution only sampled planning units that contained individuals belonging to one of the two main groups. On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups. Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target. The only difference between the two solutions is a single planning unit. By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs. 8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).

@@

One of the key advantages of this package is that it is general enough to incorporate any spatially explicit variable as an attribute space. This package can accommodate intrinsic or extrinsic variation in the feature(s). For example, adaptation processes could be secured using environmental variation [eg. @r247]. Predator-prey processes could be conserved by capturing the overlapping distributions of predators and each of their prey species [eg. @r471; @r485]. Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg. @r483]. As long as the variation can be described using Euclidean distances--which could be achieved by transforming or applying dimension reduction techniques--the \texttt{R} package can be used to obtain a representative sample.

The degree to which a prioritization truly secures a representative sample of a feature depends on the attribute spaces and the distribution of demand points chosen by the decision maker. Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points. As a consequence, if the planner uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not effectively represent the decision maker's objectives. We therefore stress that planners must carefully consider which biodiversity processes need to be reflected in the prioritization, and which spatial data can be used to represent these processes. Note that maximizing for one attribute space can result in trading off against another. For instance, to maximize the geographic spread of a prioritization, reserves need to be further away from each other, yet to maximize connectivity, reserves may need to be closer together. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space. To assist the generation of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see documentation for the `make.DemandPoints` function). These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].

The formulation requires spatially comprehensive data to map attribute spaces to planning units. However, many real-world data sets are patchy. For example, due to constrained resources, genetic data may have only been collected in some planning units. To use such patchy data with the \texttt{raptr R} package, the planning units that are missing data must either be omitted (as we did in the second case-study), or their missing values can be estimated. Missing data could be estimated using spatially explicit models [eg. krigging or generalized dissimilarity models; @r467; @r466]. This approach has been successfully applied to a range of biological data sets [eg. @r468].

>>>>

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across this study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 4a). The second species was simulated to represent a species with an idealized distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 4c). The third species represents a species with two distinct populations (Fig. 2c). This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 4d). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were set as the planing units' centroids and were weighted according to the units' probability of occupancy.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 4a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 4b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4d} \\
%
\end{align*}

We generated four solutions for each species. First to represent solutions generated using conventional methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. Second to show how the addition of representation targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets. Third to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`. Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.

@@

The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig. 3d). This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented. This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig. 3c) contained the same number of planning units as the solution with just the amount targets (Fig. 3a). Overall, these results show that under the simplest of conditions, using just amount targets results in an under-specified reserve selection problem that is unlikely to return an effective solution.

@@

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is highly fragmented.

@@

The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig. 2c). The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets prioritized individuals in just one of the two populations (Fig. 3i). While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig. 3k). However, the addition of geographic space targets resolved these issues.

@@

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63]. We selected Queensland, Australia as the study area because it contains a broad range of different habitats. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the scores from the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation; Fig. 4).

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig. S2). To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method. For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location. We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231]. We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations. These uniformly distributed points and associated density estimations used as demand point coordinates and weights (respectively).

We generated two solutions. The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount targets for all species. The second solution was generated the same amount target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.

### Results
Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches. (Figure S1). This solution secured over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`. Yet it failed to achieve this for the `r parsed.not.representative.space.held.names`. This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritization securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}) as a simple case study, because it has been found to exhibit significant genetic structure [@r462] as a simple case study. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the species was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using the grid cells' centroids and inverse great circle distances [using the \texttt{ape R} package; @r464].

The grid cells were used as planning units for generating prioritizations. The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit. Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points. To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell. We used these values to represent opportunity cost (Fig. 7b).

@@

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs. 8a and 8c). This solution only conserved individuals in one of the two main genetic groups. In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target. Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs. 8b and 8d). Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2.spp.DF$Score[2],2)` compared to `r round(cs2.spp.DF$Score[1],2)` total cost).

@@

One of the key advantages of this package is that it is general enough to incorporate almost any spatially explicit variable. This package can accommodate intrinsic or extrinsic variation in the feature(s). For example, adaptation processes could be secured using environmental variation [eg. @r247], and trophic processes could be conserved by capturing the overlapping distributions of predator and prey species [eg. @r471; @r485]. Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg. @r483]. As long as the variation can be described using Euclidean distances--which could be achieved through transformation or dimension reduction--the \texttt{R} package can be used to obtain a representative sample.

The degree to which a prioritization truly secures a representative sample of a feature depends on the attribute space(s) and the distribution of demand points chosen by the conservation planner. The optimal solution will not effectively represent the decision maker's objectives if an inapropriate set of spatial variables or demand points are used to represent the attribute space. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space. To assist the generation of appropriate demand points, the \texttt{R} package provides several routines that use the distribution of the features in spatial data sets to make demand points (see documentation for the `make.DemandPoints` function).

Attribute data need to be spatially comprehensive to map planning units to attribute spaces. However, many real-world data sets are patchy. For example, due to constrained resources, genetic data may only be available for some planning units. To use such patchy data, planners could omit the planning units that are missing data, or estimate the missing data using spatially explicit models [eg. krigging or generalized dissimilarity models; @r467; @r466]. This approach has been successfully applied to a range of biological data sets [eg. @r468].
